---
title: "ChThree_GCB_6_5_24"
output: html_document
date: "2024-06-05"
editor_options: 
  chunk_output_type: console
---

Set WD
```{r}
setwd("/Users/whiteabalone/Desktop/WACBP/gradschool/UVM/Copepods/ExperimentalData_UVM")
```

Load in libraries
```{r}
library(lsr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggExtra)
library(rcompanion)
library(FSA)
library(car)
library(ARTool)
library(rstatix)
library(reshape2)
library(ecotox)
library(MASS)
library(drc)
library(broom)
library(maps)
library(drc)
library(lme4)
library(lmtest)
library(fmsb)
library(faraway)
library(rsq)
library(ggcorrplot)
library(ggpmisc)
library(dplyr)
library(leaps)
library(AICcmodavg)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(corrplot)
library(gridExtra)
library(grid)
library(cowplot)
library(lattice)
```

# Map code for Figure One by Melissa Pespeni
# Get world map data
```{r}
world_map <- map_data("world")
```



# Sample colors for each point
#point_colors <- c("blue", "green", "red")
```{r}
point_colors_noCT <- c("dodgerblue","springgreen", "deeppink")
```


# Create a grayscale map with focused coordinates

```{r}

ggplot() +
  geom_map(data = world_map, map = world_map,
           aes(x = long, y = lat, map_id = region),
           fill = "grey", color = NA, size = 0.2) + # color NA gets rid of political boundaries
  coord_fixed(xlim = c(-85, -61),
              ylim = c(24.5, 46)) +
  theme_void() +
               
  # Add point markers based on latitude and longitude
  geom_point(data = data.frame(longitude = c(-68.54,-73.03, -82.01),
                               latitude = c(44.02, 40.95, 26.48)),
             aes(x = longitude, y = latitude), shape = 23, alpha = 0.8,
             fill = point_colors_noCT, color = "black",size = 10) +
  
  # Add labels to the points
  geom_text(data = data.frame(longitude = c(-68.54,-73.03, -82.01),
                              latitude = c(44.02,40.95, 26.48)),
                              label = c("ME","LIS", "FL"),
            aes(x = longitude, y = latitude, label = label),
            hjust = 0, vjust = 0, nudge_x = .5, nudge_y = -1.5, size = 10, fontface = "bold") +
  
  # Add a box around the plot
  theme(
    panel.border = element_rect(color = "black", size = 1, fill = NA),
    plot.margin = margin(10, 10, 10, 10)
  )
```


# Data analysis and plotting for ULT and LT50

Loading in ULT Data
```{r}

mydata <- read.table("LHS_ctmax_Complete.csv", header=TRUE, 
  	sep=",")
head(mydata)
```

Looking to see if data is normally distributed

```{r}
ggqqplot(mydata$Ctmax) 
# Q-Q plot (or quantile-quantile plot) draws the correlation between a given sample 
# and the normal distribution. A 45-degree reference line is also plotted.

ggdensity(mydata$Ctmax, xlab = "Ctmax") 

shapiro.test(mydata$Ctmax)
```
	Shapiro-Wilk normality test

data:  mydata$Ctmax
W = 0.97046, p-value = 0.02188


```{r}
hist(mydata$Ctmax, 
     main="Ctmax", 
     xlab="Ctmax", 
     border="light blue", 
     col="blue", 
     las=1, 
     breaks=5)
```

Checking for homogeneity of variance
```{r}
leveneTest(Ctmax~Population*Stage, data=mydata)
```
Levene's Test for Homogeneity of Variance (center = median)
      Df F value Pr(>F)
group  8  0.8667 0.5475
      93     
      
Aligned Rank Transform ANOVA for conducting nonparametric analyses of variance on factorial models
```{r}
mydata$Population <- as.factor(mydata$Population)

mydata$Stage <- as.factor(mydata$Stage)
```


```{r}
m <- art(Ctmax ~ Population*Stage, data=mydata)
summary(m)
```
Aligned Rank Transform of Factorial Model

Call:
art(formula = Ctmax ~ Population * Stage, data = mydata)

Column sums of aligned responses (should all be ~0):
      Population            Stage Population:Stage 
               0                0                0 

F values of ANOVAs on aligned responses not of interest (should all be ~0):
    Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
0.000000 0.000000 0.000000 0.009425 0.017514 0.033199 


```{r}
ARTaov <- anova(m)
summary(ARTaov)
ARTaov
```
Analysis of Variance of Aligned Rank Transformed Data

Table Type: Anova Table (Type III tests) 
Model: No Repeated Measures (lm)
Response: art(Ctmax)

                   Df Df.res F value     Pr(>F)    
1 Population        2     93  36.893 1.6000e-12 ***
2 Stage             2     93  26.197 9.4658e-10 ***
3 Population:Stage  4     93   1.700    0.15663 

```{r}


ARTaov$part.eta.sq = with(ARTaov, `Sum Sq`/(`Sum Sq` + `Sum Sq.res`))

ARTaov
```
                   Df Df.res F value     Pr(>F) part.eta.sq    
1 Population        2     93  36.893 1.6000e-12    0.442397 ***
2 Stage             2     93  26.197 9.4658e-10    0.360355 ***
3 Population:Stage  4     93   1.700    0.15663    0.068135 

Using the same package to look at contrasts
```{r}
ModelContrast <- art.con(m,"Stage")

ModelContrast

```
 contrast             estimate   SE df t.ratio p.value
 Adult - Copepodite      -41.6 5.94 93  -6.998  <.0001
 Adult - Nauplii         -10.4 5.80 93  -1.801  0.1748
 Copepodite - Nauplii     31.1 5.94 93   5.240  <.0001
 
 
```{r}
art.con(m,"Population")
```
 contrast          estimate   SE df t.ratio p.value
 Florida - Maine       46.6 5.57 93   8.363  <.0001
 Florida - NewYork     15.2 5.66 93   2.693  0.0227
 Maine - NewYork      -31.4 5.49 93  -5.711  <.0001
 
Code for Figure Two
```{r}
mysubdata <- within(mydata, rm(Plate))
mysubsubdata <- within(mysubdata, rm(Sex))
mysubsubdata

mysubsetdata <- within(mysubsubdata, rm(MeanMax))
mysubsetdata
```


```{r}
mydata_melt <- melt(mysubsetdata)
mydata_melt
```



Now let's try plotting again with this new data frame
```{r}
plot1 <- (ggplot(mydata_melt,aes(x=Stage,y=value,fill=Population))+geom_boxplot())

plot2 <- (plot1 + geom_dotplot(binaxis='y', stackdir='center',
                 position=position_dodge(.85))+scale_fill_manual(values=c("grey3","grey3","grey3") ))
plot2

plot2 + scale_fill_manual(values=c("red4","red3","tomato"))
```


Need to reorder graph

```{r}
mydata_melt$Stage <- factor(mydata_melt$Stage,
    levels = c("Nauplii","Copepodite","Adult"),ordered = TRUE)

mydata_melt$Population <- factor(mydata_melt$Population,
    levels = c("Maine","NewYork","Florida"),ordered = TRUE)

plot1 <- (ggplot(mydata_melt,aes(x=Stage,y=value,fill=Population))+geom_boxplot(width=0.9/length(unique(mydata_melt$Stage)),position=position_dodge(0.5)))

plot2 <- (plot1 + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75,
                 position=position_dodge(0.5))+scale_fill_manual(values=c("grey3","grey3","grey3") ))
plot2

plot2 +scale_fill_manual(values=c("dodgerblue", "springgreen", "deeppink")) +ylab("Upper Lethal Temperature (°C)") + theme_bw(base_size=25) + scale_y_continuous(breaks=seq(30,42,by=2)) 
```

For figure legends need to record number of individuals in each treatment
```{r}
count(mydata, Stage, Population)
``` 


LT50 Data analysis
```{r}
dat <- read_csv("LD50Data.csv")

dat$Stage <- factor(dat$Stage , levels = c("Nauplii", "Copepodite", "Adult"))

dat$Population <- factor(dat$Population , levels = c("Maine", "NewYork", "Florida"))
```

```{r}
fitPop <- drm(Alive/Total~Temp, Population, data=dat,type='binomial',fct=LL.3())
 
ED(fitPop, 50)

fitStage <- drm(Alive/Total~Temp, Stage, data=dat,type='binomial',fct=LL.3())
plotStage <- plot(fitStage, col=T)
ED(fitStage, 50)
```
Estimated effective doses

             Estimate Std. Error
e:Florida:50 38.09878    0.38248
e:Maine:50   35.23082    0.34569
e:NewYork:50 37.13733    0.44536

Estimated effective doses

                Estimate Std. Error
e:Adult:50      36.10212    0.40128
e:Copepodite:50 37.89030    0.51436
e:Nauplii:50    36.25248    0.63576


Code for Figure Three
```{r}
exploratory_curve <-
      ggplot(data = dat,
           aes(
             x = Temp,
             y = Alive/Total,
             # You can play aroung the group value
             
             color = Population,
             shape = Population)) +
      geom_point(size=5) +
      geom_smooth( size = 2.5,
        method = drm,
        method.args = list(
          fct = LL.3()),
        se = FALSE
      ) + scale_color_manual(values=c("dodgerblue", "springgreen", "deeppink"))+ylab('Proportion Alive')+xlab('Temperature (°C)')+   facet_grid(dat$Stage) +theme_bw() +
theme(text = element_text(size=25),legend.position="bottom")

exploratory_curve
```

Environmental Data analysis and figures


```{r}
mydata <- read.csv('LHS_ctmax_Complete_ENVDATTEMP_6_3_24.csv')
```


```{r}
#Removing non numerical variables
NumericData <- subset(mydata, select = -c(1, 2, 3, 5))
NumericData
datamatrix = cor(NumericData)
  corrplot(datamatrix, method = 'number')
  
#Want to rename the columns
  names(NumericData) <- c('ULT','DecadeMeanTemp', 'DecadeMaxTemp', 'DecadeMinTemp', 'DecadeTempRange', 'SummerMeanTemp', 'SummerMaxTemp', 'SummerMinTemp', 'SummerTempRange', 'CollectionMeanTemp', 'CollectionMaxTemp', 'CollectionMinTemp', 'CollectionTempRange')
```

```{r}
corrplot(cor(NumericData),
   method = "color", 
   addCoef.col="black", 
   order = "AOE", 
   number.cex=0.75)
```

Supplemental Figure One
```{r}

corrplot(as.matrix(cor(NumericData)), tl.cex = 0.5, tl.col = "black", method = "color", 
         outline = T,  order="hclust", 
         addCoef.col = "black", number.digits = 2, number.cex = .5, 
         cl.pos = 'b', cl.cex = .5, addrect = 3, rect.lwd = 3, 
         col = colorRampPalette(c("midnightblue", "white","darkred"))(100))
```

Plotting Environmental Data Variables against ULT for Figure Four

```{r}

#first plot decade min mean max and range

#DecadeMin

mydata$Population <- factor(mydata$Population , levels = c("Maine", "NewYork", "Florida"))

 p1 <- DecadeMin <- ggplot(mydata, aes(x=DecadeMin, y=ULT))  +
  geom_point(aes(shape=Population, color=Population), size = 5)+ 
  scale_shape_manual(values = c(0, 1, 2)) + geom_smooth(se=FALSE, method=lm)+ ylab("")  + xlab('Minimum Temperature') + theme_classic(base_size=12) +theme(legend.position = "none")+xlim(5,40)+scale_y_continuous(minor_breaks = seq(30 , 48, 1), breaks = seq(30, 48, 4), limits=c(30,48))+ scale_color_manual(values=c("dodgerblue", "springgreen", "deeppink"))+stat_cor(aes(label = ..rr.label..), color = "red", geom = "label", label.x = 10, label.y = 46, size=4.5)

#DecadeMean

p2 <- ggplot(mydata, aes(x=DecadeMean, y=ULT))  +
  geom_point(aes(shape=Population, color=Population), size = 5)+ 
  scale_shape_manual(values = c(0, 1, 2)) + geom_smooth(se=FALSE, method=lm)+ ylab("")  + xlab('Mean Temperature') + theme_classic(base_size=12)  +theme(legend.position = "none")+ xlim(5,40)+scale_y_continuous(minor_breaks = seq(30 , 48, 1), breaks = seq(30, 48, 4), limits=c(30,48))+ scale_color_manual(values=c("dodgerblue", "springgreen", "deeppink"))+stat_cor(aes(label = ..rr.label..), color = "red", geom = "label", label.x = 10, label.y = 46, size=4.5)

#DecadeMax

p3 <- ggplot(mydata, aes(x=DecadeMax, y=ULT))  +
  geom_point(aes(shape=Population, color=Population), size = 5)+ 
  scale_shape_manual(values = c(0, 1, 2)) + geom_smooth(se=FALSE, method=lm)+ ylab("")  + xlab('Maximum Temperature ') + theme_classic(base_size=12)  +theme(legend.position = "none")+ xlim(5,40)+scale_y_continuous(minor_breaks = seq(30 , 48, 1), breaks = seq(30, 48, 4), limits=c(30,48))+ scale_color_manual(values=c("dodgerblue", "springgreen", "deeppink"))+stat_cor(aes(label = ..rr.label..), color = "red", geom = "label", label.x = 10, label.y = 46, size=4.5)

#DecadeRange

p4 <- ggplot(mydata, aes(x=DecadeRange, y=ULT))  +
  geom_point(aes(shape=Population, color=Population), size = 5)+ 
  scale_shape_manual(values = c(0, 1, 2)) + geom_smooth(se=FALSE, method=lm)+ ylab("")  + xlab('Temperature Range (°C)') + theme_classic(base_size=12)  +theme(legend.position = "none")+ xlim(0,25)+scale_y_continuous(minor_breaks = seq(30 , 48, 1), breaks = seq(30, 48, 4), limits=c(30,48))+ scale_color_manual(values=c("dodgerblue", "springgreen", "deeppink"))+stat_cor(aes(label = ..rr.label..), color = "red", geom = "label", label.x = 10, label.y = 46, size=4.5)

#SummerMin

p5 <- ggplot(mydata, aes(x=SummerMin, y=ULT))  +
  geom_point(aes(shape=Population, color=Population), size = 5)+ 
  scale_shape_manual(values = c(0, 1, 2)) + geom_smooth(se=FALSE, method=lm)+ ylab("")  + xlab('Minimum Temperature') + theme_classic(base_size=12) +theme(legend.position = "none")+ xlim(5,40)+scale_y_continuous(minor_breaks = seq(30 , 48, 1), breaks = seq(30, 48, 4), limits=c(30,48))+ scale_color_manual(values=c("dodgerblue", "springgreen", "deeppink"))+stat_cor(aes(label = ..rr.label..), color = "red", geom = "label", label.x = 10, label.y = 46, size=4.5)

#SummerMean

p6 <- ggplot(mydata, aes(x=SummerMean, y=ULT))  +
  geom_point(aes(shape=Population, color=Population), size = 5)+ 
  scale_shape_manual(values = c(0, 1, 2)) + geom_smooth(se=FALSE, method=lm)+ ylab("")  + xlab('Mean Temperature ') + theme_classic(base_size=12)  +theme(legend.position = "none")+ xlim(5,40)+scale_y_continuous(minor_breaks = seq(30 , 48, 1), breaks = seq(30, 48, 4), limits=c(30,48))+ scale_color_manual(values=c("dodgerblue", "springgreen", "deeppink"))+stat_cor(aes(label = ..rr.label..), color = "red", geom = "label", label.x = 10, label.y = 46, size=4.5)

#SummerMax

p7 <- ggplot(mydata, aes(x=SummerMax, y=ULT))  +
  geom_point(aes(shape=Population, color=Population), size = 5)+ 
  scale_shape_manual(values = c(0, 1, 2)) + geom_smooth(se=FALSE, method=lm)+ ylab("")  + xlab('Maximum Temperature') + theme_classic(base_size=12)  +theme(legend.position = "none")+ xlim(5,40)+scale_y_continuous(minor_breaks = seq(30 , 48, 1), breaks = seq(30, 48, 4), limits=c(30,48))+ scale_color_manual(values=c("dodgerblue", "springgreen", "deeppink"))+stat_cor(aes(label = ..rr.label..), color = "red", geom = "label", label.x = 10, label.y = 46, size=4.5)

#SummerRange

p8 <- ggplot(mydata, aes(x=SummerRange, y=ULT))  +
  geom_point(aes(shape=Population, color=Population), size = 5)+ 
  scale_shape_manual(values = c(0, 1, 2)) + geom_smooth(se=FALSE, method=lm)+ ylab("")  + xlab('Temperature Range (°C)') + theme_classic(base_size=12) +theme(legend.position  = "none")+ xlim(0,25)+scale_y_continuous(minor_breaks = seq(30 , 48, 1), breaks = seq(30, 48, 4), limits=c(30,48))+ scale_color_manual(values=c("dodgerblue", "springgreen", "deeppink"))+stat_cor(aes(label = ..rr.label..), color = "red", geom = "label", label.x = 10, label.y = 46, size=4.5)

#CollectionMin

p9 <- ggplot(mydata, aes(x=CollectionMin, y=ULT))  +
  geom_point(aes(shape=Population, color=Population), size = 5)+ 
  scale_shape_manual(values = c(0, 1, 2)) + geom_smooth(se=FALSE, method=lm)+ ylab("")  + xlab('Minimum Temperature') + theme_classic(base_size=12)  +theme(legend.position = "none")+ xlim(5,40) +scale_y_continuous(minor_breaks = seq(30 , 48, 1), breaks = seq(30, 48, 4), limits=c(30,48))+ scale_color_manual(values=c("dodgerblue", "springgreen", "deeppink"))+stat_cor(aes(label = ..rr.label..), color = "red", geom = "label", label.x = 10, label.y = 46, size=4.5)

#CollectionMean

p10 <- ggplot(mydata, aes(x=CollectionMean, y=ULT))  +
  geom_point(aes(shape=Population, color=Population), size = 5)+ 
  scale_shape_manual(values = c(0, 1, 2)) + geom_smooth(se=FALSE, method=lm)+ ylab("")  + xlab('Mean Temperature') + theme_classic(base_size=12)  +theme(legend.position = "none")+ xlim(5,40)+scale_y_continuous(minor_breaks = seq(30 , 48, 1), breaks = seq(30, 48, 4), limits=c(30,48))+ scale_color_manual(values=c("dodgerblue", "springgreen", "deeppink"))+stat_cor(aes(label = ..rr.label..), color = "red", geom = "label", label.x = 10, label.y = 46, size=4.5)

#CollectionMax

p11 <- ggplot(mydata, aes(x=CollectionMax, y=ULT))  +
  geom_point(aes(shape=Population, color=Population), size = 5)+ 
  scale_shape_manual(values = c(0, 1, 2)) + geom_smooth(se=FALSE, method=lm)+ ylab("")  + xlab('Maximum Temperature') + theme_classic(base_size=12) +theme(legend.position = "none")+ xlim(5,40)+scale_y_continuous(minor_breaks = seq(30 , 48, 1), breaks = seq(30, 48, 4), limits=c(30,48))+ scale_color_manual(values=c("dodgerblue", "springgreen", "deeppink"))+stat_cor(aes(label = ..rr.label..), color = "red", geom = "label", label.x = 10, label.y = 46, size=4.5)

#CollectionRange

p12 <- ggplot(mydata, aes(x=CollectionRange, y=ULT))  +
  geom_point(aes(shape=Population, color=Population), size = 5)+ 
  scale_shape_manual(values = c(0, 1, 2)) + geom_smooth(se=FALSE, method=lm)+ ylab("")  + xlab('Temperature Range (°C)') + theme_classic(base_size=12) +theme(legend.position = "none")+ xlim(0,25)+scale_y_continuous(minor_breaks = seq(30 , 48, 1), breaks = seq(30, 48, 4), limits=c(30,48))+ scale_color_manual(values=c("dodgerblue", "springgreen", "deeppink"))+stat_cor(aes(label = ..rr.label..), color = "red", geom = "label", label.x = 10, label.y = 46, size=4.5)

#for adding r2 if we want to add back in at some point
#+stat_cor(aes(label = ..r.label..), color = "red", geom = "label", label.x = -8, label.y = 45, size=5)

```

```{r}
title1=text_grob('Upper Lethal Temperature (°C)', size = 20, rot=90)
title2=text_grob('Decade', size = 20)
```

Put plots together
```{r}
myLegend <- get_legend(p1 + theme(legend.position="bottom") )

grid.arrange(arrangeGrob(p1,p2,p3,p4,top=textGrob("Decade", gp=gpar(fontsize=15,font=8)), ncol=4), 
             arrangeGrob(p5,p6,p7,p8,top=textGrob("Summer", gp=gpar(fontsize=15,font=8)), ncol=4), 
             arrangeGrob(p9,p10,p11,p12,top=textGrob("Month", gp=gpar(fontsize=15,font=8)), ncol=4), 
             myLegend,
             nrow=4, heights=c(1,1,1,0.2),
             left=title1)

```

Environmental Data analysis

Building individual models to compare
```{r}
ModelDecadeMean <- lm(ULT~DecadeMean, data = mydata)
ModelDecadeMax <- lm(ULT~DecadeMax, data = mydata)
ModelDecadeMin <- lm(ULT~DecadeMin, data = mydata)
ModelDecadeRange <- lm(ULT~DecadeRange, data = mydata)
ModelSummerMean <- lm(ULT~SummerMean, data = mydata)
ModelSummerMax <- lm(ULT~SummerMax, data = mydata)
ModelSummerMin <- lm(ULT~SummerMin, data = mydata)
ModelSummerRange <- lm(ULT~SummerRange, data = mydata)
ModelCollectionMean <- lm(ULT~CollectionMean, data = mydata)
ModelCollectionMax <- lm(ULT~CollectionMax, data = mydata)
ModelCollectionMin <- lm(ULT~CollectionMin, data = mydata)
ModelCollectionRange <- lm(ULT~CollectionRange, data = mydata)
```

```{r}
#define list of models
models <- list(ModelDecadeMean, ModelDecadeMax, ModelDecadeMin, ModelDecadeRange, ModelSummerMean, ModelSummerMax, ModelSummerMin, ModelSummerRange, ModelCollectionMean, ModelCollectionMax, ModelCollectionMin, ModelCollectionRange)

#specify model names
mod.names <- c('DMean', 'DMax', 'DMin', 'DRange', 'SMean', 'SMax', 'SMin', 'SRange', 'CMean', 'CMax', 'CMin', 'CRange')

#calculate AIC of each model
aictab(cand.set = models, modnames = mod.names)
```
Model selection based on AICc:

       K   AICc Delta_AICc AICcWt Cum.Wt      LL
SMax   3 413.95       0.00   0.75   0.75 -203.85
SMean  3 417.53       3.57   0.12   0.87 -205.64
DMean  3 418.11       4.16   0.09   0.96 -205.93
CMax   3 421.18       7.22   0.02   0.98 -207.47
SMin   3 422.80       8.85   0.01   0.99 -208.28
DMax   3 425.24      11.29   0.00   0.99 -209.50
DMin   3 426.02      12.07   0.00   1.00 -209.89
DRange 3 426.55      12.60   0.00   1.00 -210.15
CMean  3 426.66      12.71   0.00   1.00 -210.21
CMin   3 427.72      13.77   0.00   1.00 -210.74
SRange 3 439.32      25.37   0.00   1.00 -216.54
CRange 3 439.45      25.50   0.00   1.00 -216.60

Lower Lethal Salinity Analysis and Plots

Loading in LLS Data
```{r}
Saldata <- read.table("LLSDataJess.csv", header=TRUE, 
  	sep=",")
head(Saldata)

names(Saldata)<- c("Population","Stage","LLS")

Saldata
```

For figure legends need to record number of individuals in each treatment
```{r}
count(Saldata, Stage, Population)
```

Looking to see if data is normally distributed

```{r}
ggqqplot(Saldata$LLS) 
# Q-Q plot (or quantile-quantile plot) draws the correlation between a given sample 
# and the normal distribution. A 45-degree reference line is also plotted.

ggdensity(Saldata$LLS, xlab = "LLS") 

shapiro.test(Saldata$LLS)
```

data:  mydata$LLS
W = 0.87258, p-value = 5.042e-08
Super not normal 

```{r}
hist(Saldata$LLS, 
     main="LLS", 
     xlab="LLS", 
     border="light blue", 
     col="blue", 
     las=1, 
     breaks=5)
```
Data skewed right


```{r}
str(Saldata)
```

105 observations


Checking for homogeneity of variance
```{r}
leveneTest(LLS~Population*Stage, data=Saldata)
```
      Df F value  Pr(>F)  
group  8  2.2525 0.02992 *

Not homogeneous

So we use ART ANOVA
```{r}
Saldata$Population <- as.factor(Saldata$Population)

Saldata$Stage <- as.factor(Saldata$Stage)

Saldata
```

```{r}

mawdel <- art(LLS ~ Population*Stage, data=Saldata)
summary(mawdel)
```
Call:
art(formula = Ctmax ~ Population * Stage, data = mydata)

Column sums of aligned responses (should all be ~0):
      Population            Stage Population:Stage 
               0                0                0 

F values of ANOVAs on aligned responses not of interest (should all be ~0):
    Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
0.000000 0.000000 0.000000 0.009425 0.017514 0.033199 
Warning message:
In summary.art(m) :
  F values of ANOVAs on aligned responses not of interest are not all ~0. ART may not be appropriate.
  


```{r}
thing <- anova(mawdel)
summary(thing)
thing

thing$part.eta.sq = with(thing, `Sum Sq`/(`Sum Sq` + `Sum Sq.res`))

thing

```
                   Df Df.res F value     Pr(>F)    
1 Population        2     96  4.0254   0.020952   *
2 Stage             2     96 17.3716 3.6384e-07 ***
3 Population:Stage  4     96  7.5622 2.4382e-05 ***

Using the same package to look at contrasts
```{r}
ModelContrast <- art.con(mawdel,"Stage")

ModelContrast

```
contrast              estimate   SE df t.ratio p.value
 Adults - Copepodites   -32.869 6.50 96  -5.058  <.0001
 Adults - Nauplii         0.591 6.40 96   0.092  0.9953
 Copepodites - Nauplii   33.460 6.45 96   5.186  <.0001
 
Calculating Cohen's d as a measure of effect size 
```{r}
library(emmeans)
m.art.Stage = artlm(mawdel, "Stage")
Stage.contrasts.art = summary(pairs(emmeans(m.art.Stage, ~ Stage)))

Stage.contrasts.art$d = Stage.contrasts.art$estimate / sigmaHat(m.art.Stage)
Stage.contrasts.art
```
  contrast              estimate   SE df t.ratio p.value       d
 Adults - Copepodites   -32.869 6.50 96  -5.058  <.0001 -1.2189
 Adults - Nauplii         0.591 6.40 96   0.092  0.9953  0.0219
 Copepodites - Nauplii   33.460 6.45 96   5.186  <.0001  1.2408
 
```{r}
art.con(mawdel,"Population")
```
 contrast    estimate   SE df t.ratio p.value
 HI - PGH       17.94 7.08 96   2.536  0.0340
 HI - UConn      1.43 7.13 96   0.201  0.9779
 PGH - UConn   -16.51 7.02 96  -2.350  0.0538





Now let's try plotting again with this new data frame
```{r}
Saldata
SalDataNoEnvDat <- Saldata[, c("Population", "Stage", "LLS")]

Saldata_melt <- melt(SalDataNoEnvDat)
Saldata_melt

plot1 <- (ggplot(Saldata_melt,aes(x=Stage,y=value,fill=Population))+geom_boxplot())

plot2 <- (plot1 + geom_dotplot(binaxis='y', stackdir='center',
                 position=position_dodge(.85))+scale_fill_manual(values=c("grey3","grey3","grey3") ))
plot2

plot2 + scale_fill_manual(values=c("grey96","grey48","grey12"), labels=c("Maine","Connecticut","Florida"))
```


Need to reorder graph

```{r}
 Saldata_melt$Stage <- factor(Saldata_melt$Stage,
     levels = c("Nauplii","Copepodites","Adults"),ordered = TRUE)
 
 Saldata_melt$Population <- factor(Saldata_melt$Population,
     levels = c("HI","UConn","PGH"),ordered = TRUE)
 Saldata_melt

plot1 <- (ggplot(Saldata_melt,aes(x=Stage,y=value,fill=Population))+geom_boxplot(width=0.9/length(unique(Saldata_melt$Stage)),position=position_dodge(.5)))

plot2 <- (plot1 + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75,
                 position=position_dodge(.5))+scale_fill_manual(values=c("grey3","grey3","grey3") ))
plot2

plot2 + scale_fill_manual(values=c("dodgerblue", "springgreen", "deeppink"),labels=c("Maine","Connecticut","Florida")) +ylab("Lower Lethal Salinity (ppt)") + theme_bw(base_size=25) + scale_y_continuous(breaks=seq(0,10,by=2))


```


