---
title: "ThermalTracking_OikosResubmission"
output: html_document
date: "2025-06-12"
editor_options: 
  chunk_output_type: console
---

Set WD
```{r}
setwd("/Users/whiteabalone/Desktop/WACBP/gradschool/UVM/Copepods/ExperimentalData_UVM")
```


Load in libraries
```{r}
library(lsr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggExtra)
library(rcompanion)
library(FSA)
library(car)
library(ARTool)
library(rstatix)
library(reshape2)
library(ecotox)
library(MASS)
library(drc)
library(broom)
library(maps)
library(drc)
library(lme4)
library(lmtest)
library(fmsb)
library(faraway)
library(rsq)
library(ggcorrplot)
library(ggpmisc)
library(leaps)
library(AICcmodavg)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(corrplot)
library(gridExtra)
library(grid)
library(cowplot)
library(lattice)
library(gtable)
library(mgcv)
library(olsrr)
library(emmeans)
library(Rfit)
```

# Map code for Figure One by Melissa Pespeni
# Get world map data
```{r}
world_map <- map_data("world")
```

```{r}
east_coast_map <- world_map %>%
  filter(region == "USA") %>%
  filter(long >= -85 & long <= -61 & lat >= 24.5 & lat <= 46)
```


# Sample colors for each point

```{r}
point_colors <- c("dodgerblue", "springgreen", "springgreen4", "deeppink")
```

# Create a grayscale map with focused coordinates

```{r}

ggplot() +
  geom_map(data = east_coast_map, map = world_map,
           aes(x = long, y = lat, map_id = region),
           fill = "grey", color = NA, size = 0.2) + # color NA gets rid of political boundaries

  coord_fixed(xlim = c(-85, -61),
              ylim = c(24.5, 46)) +
  theme_void() +

  
  # Add point markers based on latitude and longitude
  geom_point(data = data.frame(longitude = c(-68.54,-72,-73.03, -82.01),
                               latitude = c(44.02, 41.32, 40.95, 26.48)),
             aes(x = longitude, y = latitude), shape = 23, alpha = 0.8,
             fill = point_colors, color = "black",size = 6) +
  
  # Add labels to the points
  geom_text(data = data.frame(longitude = c(-68.54,-72,-72, -82.01),
                              latitude = c(44.02, 41.5, 40.5, 26.48)),
                              label = c("ME","CT","NY", "FL"),
            aes(x = longitude, y = latitude, label = label),
            hjust = 0, vjust = 0, nudge_x = .5, nudge_y = -1.5, size = 6, fontface = "bold") +
  
  # Add a box around the plot
  theme(
    panel.border = element_rect(color = "black", size = 1, fill = NA),
    plot.margin = margin(10, 10, 10, 10)
  ) +
    annotate("text", x = -66, y = 40, label = "      } Long Island Sound", size = 7, color = "black", fontface = "bold")
```

# Data analysis and plotting for ULT and LT50

Loading in ULT Data
```{r}

mydata <- read.table("LHS_ULT_OikosResubmission.csv", header=TRUE, 
  	sep=",")
head(mydata)
```

NY mean and se
```{r}
mean(mydata$ULT[mydata$CollectionLocation=="NewYork" & mydata$Stage=="Copepodite"])

standard_error <- sd(mydata$ULT[mydata$CollectionLocation=="NewYork" & mydata$Stage=="Copepodite"]) / sqrt(length(mydata$ULT[mydata$CollectionLocation=="NewYork" & mydata$Stage=="Copepodite"]))

standard_error
```

Run linear model and look at residuals to test normality assumptions
```{r}
ModULT<- lm(ULT~CollectionLocation+Stage, data=mydata)
summary(ModULT)
plot(ModULT)
ols_test_normality(ModULT)

```


Aligned Rank Transform ANOVA for conducting nonparametric analyses of variance on factorial models
```{r}
mydata$CollectionLocation <- as.factor(mydata$CollectionLocation)

mydata$Stage <- as.factor(mydata$Stage)
```


```{r}
m <- art(ULT ~ CollectionLocation*Stage, data=mydata)
summary(m)
```


```{r}

library(lmerTest)  # this must come first
library(ARTool)
ARTaov <- anova(m)
summary(ARTaov)
ARTaov
```


```{r}
ARTaov$part.eta.sq = with(ARTaov, `Sum Sq`/(`Sum Sq` + `Sum Sq.res`))

ARTaov
```
                   

Using the same package to look at contrasts
```{r}

ModelContrast <- art.con(m,"Stage")

ModelContrast

```

 
 
```{r}
art.con(m,"CollectionLocation")
```


Code for Figure Two
```{r}
mysubdata <- within(mydata, rm(Plate))
mysubsetdata <- within(mysubdata, rm(Sex))

mysubsetdata
```


```{r}
mydata_melt <- melt(mysubsetdata)
mydata_melt
```



Now let's try plotting again with this new data frame
```{r}
plot1 <- (ggplot(mydata_melt,aes(x=Stage,y=value,fill=CollectionLocation))+geom_boxplot())

plot2 <- (plot1 + geom_dotplot(binaxis='y', stackdir='center',
                 position=position_dodge(.85))+scale_fill_manual(values=c("grey3","grey3","grey3") ))
plot2

plot2 + scale_fill_manual(values=c("red4","red3","tomato"))
```


Need to reorder graph

```{r}
mydata_melt$Stage <- factor(mydata_melt$Stage,
    levels = c("Nauplii","Copepodite","Adult"),ordered = TRUE)

mydata_melt$CollectionLocation <- factor(mydata_melt$CollectionLocation,
    levels = c("Maine","NewYork","Florida"),ordered = TRUE)

plot1 <- (ggplot(mydata_melt,aes(x=Stage,y=value,fill=CollectionLocation))+geom_boxplot(width=0.9/length(unique(mydata_melt$Stage)),position=position_dodge(0.5)))

plot2 <- (plot1 + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75,
                 position=position_dodge(0.5))+scale_fill_manual(values=c("grey3","grey3","grey3") ))
plot2

plot2 +scale_fill_manual(values=c("dodgerblue", "springgreen4", "deeppink")) +ylab("Upper Lethal Temperature (°C)") + theme_bw(base_size=25) + scale_y_continuous(breaks=seq(30,42,by=2)) 
```

For figure legends need to record number of individuals in each treatment
```{r}
count(mydata, Stage, CollectionLocation)
``` 

LT50 Data analysis
```{r}
dat <- read_csv("LD50Data.csv")

dat$Stage <- factor(dat$Stage , levels = c("Nauplii", "Copepodite", "Adult"))

dat$CollectionLocation <- factor(dat$CollectionLocation , levels = c("Maine", "NewYork", "Florida"))
```

```{r}
fitPop <- drm(Alive/Total~Temp, CollectionLocation, data=dat,type='binomial',fct=LL.3())
 
ED(fitPop, 50)

fitStage <- drm(Alive/Total~Temp, Stage, data=dat,type='binomial',fct=LL.3())
plotStage <- plot(fitStage, col=T)

ED(fitStage, 50)
```


Code for Figure Three
```{r}
exploratory_curve <-
      ggplot(data = dat,
           aes(
             x = Temp,
             y = Alive/Total,
             # You can play aroung the group value
             
             color = CollectionLocation,
             shape = CollectionLocation)) +
      geom_point(size=5) +
      geom_smooth( size = 2.5,
        method = drm,
        method.args = list(
          fct = LL.3()),
        se = FALSE
      ) + scale_color_manual(values=c("dodgerblue", "springgreen4", "deeppink"))+ylab('Proportion Alive')+xlab('Temperature (°C)')+   facet_grid(dat$Stage) +theme_bw() +
theme(text = element_text(size=25),legend.position="bottom")

exploratory_curve
```

Lower Lethal Salinity Analysis and Plots

Loading in LLS Data
```{r}

Saldata <- read.table("LLSData_OikosResubmission.csv", header=TRUE, 
  	sep=",")
head(Saldata)

```

For figure legends need to record number of individuals in each treatment
```{r}
count(Saldata, Stage, CollectionLocation)
```

```{r}
head(Saldata)
Saldata$CollectionLocation <- factor(Saldata$CollectionLocation)
Saldata$Population <- relevel(Saldata$CollectionLocation, ref = 'HI')
Saldata$Stage <- factor(Saldata$Stage)
Saldata$Stage <- relevel(Saldata$Stage, ref = 'Adults')
```

```{r}
ModLLS<- lm(LLS~CollectionLocation+Stage, data=Saldata)
summary(ModLLS)

```


```{r}
plot(ModLLS)
ols_test_normality(ModLLS)
```


ART ANOVA
```{r}
Saldata$CollectionLocation <- as.factor(Saldata$CollectionLocation)

Saldata$Stage <- as.factor(Saldata$Stage)

Saldata
```

```{r}

mawdel <- art(LLS ~ CollectionLocation*Stage, data=Saldata)
summary(mawdel)
```


```{r}
thing <- anova(mawdel)
summary(thing)
thing

thing$part.eta.sq = with(thing, `Sum Sq`/(`Sum Sq` + `Sum Sq.res`))

thing

```
                   

Using the same package to look at contrasts
```{r}
ModelContrast <- art.con(mawdel,"Stage")

ModelContrast

```

 
Calculating Cohen's d as a measure of effect size 
```{r}
m.art.Stage = artlm(mawdel, "Stage")
Stage.contrasts.art = summary(pairs(emmeans(m.art.Stage, ~ Stage)))

Stage.contrasts.art$d = Stage.contrasts.art$estimate / sigmaHat(m.art.Stage)
Stage.contrasts.art
```

 
```{r}
art.con(mawdel,"CollectionLocation")
```



Now let's try plotting again with this new data frame
```{r}
Saldata

Saldata_melt <- melt(Saldata)
Saldata_melt

plot1 <- (ggplot(Saldata_melt,aes(x=Stage,y=value,fill=CollectionLocation))+geom_boxplot())

plot2 <- (plot1 + geom_dotplot(binaxis='y', stackdir='center',
                 position=position_dodge(.85))+scale_fill_manual(values=c("grey3","grey3","grey3") ))
plot2

plot2 + scale_fill_manual(values=c("grey96","grey48","grey12"), labels=c("Maine","Connecticut","Florida"))
```


Need to reorder graph

```{r}
 Saldata_melt$Stage <- factor(Saldata_melt$Stage,
     levels = c("Nauplii","Copepodites","Adults"),ordered = TRUE)
 
 Saldata_melt$CollectionLocation <- factor(Saldata_melt$CollectionLocation,
     levels = c("HI","UConn","PGH"),ordered = TRUE)
 Saldata_melt

plot1 <- (ggplot(Saldata_melt,aes(x=Stage,y=value,fill=CollectionLocation))+geom_boxplot(width=0.9/length(unique(Saldata_melt$Stage)),position=position_dodge(.5)))

plot2 <- (plot1 + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75,
                 position=position_dodge(.5))+scale_fill_manual(values=c("grey3","grey3","grey3") ))
plot2

plot2 + scale_fill_manual(values=c("dodgerblue", "springgreen", "deeppink"),labels=c("Maine","Connecticut","Florida")) +ylab("Lower Lethal Salinity (ppt)") + theme_bw(base_size=25) + scale_y_continuous(breaks=seq(0,10,by=2))


```

iButton data plots 
```{r}
MENYNaup <- read.table("MENYNaupDat.csv", header=TRUE, 
  	sep=",")
head(MENYNaup)

MENYNAdults <- read.table("MENYAdultsDat.csv", header=TRUE, 
  	sep=",")
head(MENYNAdults)

FLAll <- read.table("FloridaAllOneDat.csv", header=TRUE, 
  	sep=",")
head(FLAll)

```

Now I want to plot each dataframe Temp by Time
```{r}
ggplot(MENYNaup, aes(x = Minutes.from.start.of.assay, y = Temp)) +
  geom_line() +
  labs(x = "Minutes after start of assay", y = "Temperature (°C)", title = "Temperature vs. Time Maine and New York nauplii") +
  theme_minimal()
```

```{r}
ggplot(MENYNAdults, aes(x = Time.after.start.of.assay, y = Temp)) +
  geom_line() +
  labs(x = "Minutes after start of assay", y = "Temperature (°C)", title = "Temperature vs. Time Maine and New York adults") +
  theme_minimal()
```

```{r}
ggplot(FLAll, aes(x = Time.after.start.of.assay, y = Temp)) +
  geom_line() +
  labs(x = "Minutes after start of assay", y = "Temperature (°C)", title = "Temperature vs. Time Florida all stages") +
  theme_minimal()
```

